---
title: "Galapagos_SH&BT_Final"
author: "Savannah Ryburn"
date: "2025-12-18"
output: html_document
---

Loading packages
```{r}
library(phyloseq)
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyverse)
library(vegan)
library(tidyr)
library(car)
library(rstatix)
library(devtools)
library(pairwiseAdonis)
library(spaa)
library(dietr)
library(ggridges)

#install.packages("ggraph") # for taxatree_plots()
#install.packages("DT") # for tax_fix_interactive()
#install.packages("corncob") # for example datasets and beta binomial models
#if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
#BiocManager::install(c("phyloseq", "microbiome", "ComplexHeatmap"), update = FALSE)

#install.packages(
#  "microViz",
#  repos = c(davidbarnett = "https://david-barnett.r-universe.dev", getOption("repos"))
#)
library(microViz)
library(ggraph)
library(DT)
library(corncob)

#install.packages("remotes")
#remotes::install_github("vmikk/metagMisc")
library(metagMisc)
```

Load phyloseq object of just fecal swabs made in script 10
```{r}
Gal_Shark_Diet.ps<-readRDS(file = "/Users/savannahryburn/Desktop/Galapagos_Sharks/10_Phyloseq/Gal_Shark_Diet.ps.RDS")

Gal_Shark_Diet.ps
```

Combine OTUs from the two fish species that weren't present in the Galapagos with the lowest taxonomic group
```{r}
# New phyloseq object so I don't mess up the old one
Gal_Shark_Diet.ps.new1 <- Gal_Shark_Diet.ps

# Extract matrix safely
otu_mat <- as(otu_table(Gal_Shark_Diet.ps.new1), "matrix")

# Double-check the OTU names (Make sure this returns exactly: "Abudefduf genus 1"  "Abudefduf taurus 1")
grep("Abudefduf", rownames(otu_mat), value = TRUE)

# save them explicitly:
otu1 <- "Abudefduf genus 1"
otu2 <- "Abudefduf taurus 1"

# Sum the two OTUs into a new named row
# Check if both OTUs exist
if (all(c(otu1, otu2) %in% rownames(otu_mat))) {
  new_otu_row <- otu_mat[otu1, ] + otu_mat[otu2, ]
  # Add the new row
  otu_mat <- rbind(otu_mat, "Abudefduf" = new_otu_row)
  # Remove the original two
  otu_mat <- otu_mat[!rownames(otu_mat) %in% c(otu1, otu2), ]
} else {
  stop("One or both OTU names not found!")
}

nrow(otu_mat)  # should be 32

# Rebuild OTU table (make sure it has the right orientation)
otu_tab_new <- otu_table(otu_mat, taxa_are_rows = TRUE)

# Replace the OTU table in your phyloseq object
Gal_Shark_Diet.ps.new1 <- phyloseq(
  otu_tab_new,
  tax_table(Gal_Shark_Diet.ps.new1),
  sample_data(Gal_Shark_Diet.ps.new1)
)

dim(otu_table(Gal_Shark_Diet.ps.new1))  # Should show 32 rows
rownames(otu_table(Gal_Shark_Diet.ps.new1))  # Should include "Abudefduf"




head(tax_table(Gal_Shark_Diet.ps.new1))

# Manually create new row
# Extract the existing tax table
tax_mat <- as(tax_table(Gal_Shark_Diet.ps.new1), "matrix")

# Create a new row (fill in with correct levels from your data structure)
new_tax_row <- c(
  Kingdom = "Eukaryota",
  Phylum = "Chordata",
  Class = "Actinopterygii",
  Order = "Perciformes",
  Family = "Pomacentridae",
  Genus = "Abudefduf",
  Species = "Abudefduf genus"  # or "Abudefduf spp." or leave NA
)

# Add the new row with the OTU name "Abudefduf"
tax_mat <- rbind(tax_mat, "Abudefduf" = new_tax_row)

# Recreate tax_table object
new_tax_table <- tax_table(tax_mat)

# Rebuild phyloseq object
Gal_Shark_Diet.ps.new1 <- phyloseq(
  otu_tab_new,
  new_tax_table,
  sample_data(Gal_Shark_Diet.ps.new1)
)

dim(otu_table(Gal_Shark_Diet.ps.new1))       # Should now be 32 x 172
rownames(otu_table(Gal_Shark_Diet.ps.new1))  # Should include "Abudefduf"
tax_table(Gal_Shark_Diet.ps.new1)["Abudefduf", ]  # Should show your new taxonomy

View(Gal_Shark_Diet.ps.new1)
View(Gal_Shark_Diet.ps.new1@tax_table)



####### Now do this again for "Cyprinodon variegatus 1" and "Actinopteri class 1"

# save them explicitly:
otu3 <- "Cyprinodon variegatus 1"
otu4 <- "Actinopteri class 1"

# Sum the two OTUs into a new named row
# Check if both OTUs exist
if (all(c(otu3, otu4) %in% rownames(otu_mat))) {
  new_otu_row <- otu_mat[otu3, ] + otu_mat[otu4, ]
  # Add the new row
  otu_mat <- rbind(otu_mat, "Actinopteri" = new_otu_row)
  # Remove the original two
  otu_mat <- otu_mat[!rownames(otu_mat) %in% c(otu3, otu4), ]
} else {
  stop("One or both OTU names not found!")
}

nrow(otu_mat)  # should be 31

# Rebuild OTU table (make sure it has the right orientation)
otu_tab_new <- otu_table(otu_mat, taxa_are_rows = TRUE)

# Replace the OTU table in your phyloseq object
Gal_Shark_Diet.ps.new1 <- phyloseq(
  otu_tab_new,
  tax_table(Gal_Shark_Diet.ps.new1),
  sample_data(Gal_Shark_Diet.ps.new1)
)

dim(otu_table(Gal_Shark_Diet.ps.new1))  # Should show 32 rows
rownames(otu_table(Gal_Shark_Diet.ps.new1))  # Should include "Abudefduf"


head(tax_table(Gal_Shark_Diet.ps.new1))

# Manually create new row
# Extract the existing tax table
tax_mat <- as(tax_table(Gal_Shark_Diet.ps.new1), "matrix")

# Create a new row (fill in with correct levels from your data structure)
new_tax_row <- c(
  Kingdom = "Eukaryota",
  Phylum = "Chordata",
  Class = "Actinopteri",
  Order = "Actinopteri class",
  Family = "Actinopteri class",
  Genus = "Actinopteri class",
  Species = "Actinopteri class"
)

# Add the new row with the OTU name "Actinopteri"
tax_mat <- rbind(tax_mat, "Actinopteri" = new_tax_row)

# Recreate tax_table object
new_tax_table <- tax_table(tax_mat)

# Rebuild phyloseq object
Gal_Shark_Diet.ps.new1 <- phyloseq(
  otu_tab_new,
  new_tax_table,
  sample_data(Gal_Shark_Diet.ps.new1)
)

dim(otu_table(Gal_Shark_Diet.ps.new1))       # Should now be 31 x 172
rownames(otu_table(Gal_Shark_Diet.ps.new1))  # Should include "Actinopteri"
tax_table(Gal_Shark_Diet.ps.new1)["Actinopteri", ]  # Should show your new taxonomy

View(Gal_Shark_Diet.ps.new1)
View(Gal_Shark_Diet.ps.new1@tax_table)
View(Gal_Shark_Diet.ps.new1@otu_table)
```

Divide blacktips and scalloped hammerheads
```{r}
Gal_Shark_Diet_SH.ps<-subset_samples(Gal_Shark_Diet.ps.new1,Species..scientific.name.=="Sphyrna lewini")
View(Gal_Shark_Diet_SH.ps)

Gal_Shark_Diet_BT.ps<-subset_samples(Gal_Shark_Diet.ps.new1,Species..scientific.name.=="Carcharhinus limbatus")
View(Gal_Shark_Diet_BT.ps)
```

Filter out host DNA
```{r}
##Scalloped Hammerheads
tax_table(Gal_Shark_Diet_SH.ps)

# Exclude "Carcharhinus genus"
Gal_Shark_Diet_SH.ps <- subset_taxa(Gal_Shark_Diet_SH.ps, species != "Sphyrna lewini")

View(Gal_Shark_Diet_SH.ps)


##Blacktips
tax_table(Gal_Shark_Diet_BT.ps)

# Exclude "Carcharhinus genus"
Gal_Shark_Diet_BT.ps <- subset_taxa(Gal_Shark_Diet_BT.ps, species != "Carcharhinus genus")

View(Gal_Shark_Diet_BT.ps)
```

Divide Blacktips by Sites (only include Puerto Grande and Cartago Bay)
```{r}
Gal_Shark_Diet_BT_PGCB.ps <- subset_samples(Gal_Shark_Diet_BT.ps, Site.Name %in% c("Puerto Grande", "Cartago Bay"))
View(Gal_Shark_Diet_BT_PGCB.ps)
```

Combine phyloseq objects for SH and BT (blacktips only from PG and CB)
```{r}
Combined_Gal_Shark_Diet.ps <- merge_phyloseq(Gal_Shark_Diet_SH.ps, Gal_Shark_Diet_BT_PGCB.ps)
```

Remove all samples with no results (blacktips only from PG and CB)
```{r}
samples_to_remove <- c("con_G15", "con_G31", "con_G43", "G105", "G106", "G109", "G116", "G117", "G121", "G125", "G127", "G128", "G129", "G134", "G163", "G166", "G190", "G082", "G083", "G084", "G086", "G088", "G090", "G092", "G093", "G095", "G096", "G097", "G143", "G151", "G153", "G154", "G156", "G158", "G168", "G173", "G177", "G178", "G180", "G182", "G184", "G185")

Combined_Gal_Shark_Diet_filtered.ps <- prune_samples(!(sample_names(Combined_Gal_Shark_Diet.ps) %in% samples_to_remove), Combined_Gal_Shark_Diet.ps)
```

Divide Blacktips by Sites (only include Puerto Grande, Cartago Bay, La Seca, and Grimanesa)
```{r}
Gal_Shark_Diet_BT_PGCBLSGR.ps <- subset_samples(Gal_Shark_Diet_BT.ps, Site.Name %in% c("Puerto Grande", "Cartago Bay", "La Seca", "Grimanesa"))
View(Gal_Shark_Diet_BT_PGCBLSGR.ps)
```

Combine phyloseq objects for SH and BT (blacktips from PG, CB, LS, and GR)
```{r}
Combined_Gal_Shark_Diet_PGCBLSGR.ps <- merge_phyloseq(Gal_Shark_Diet_SH.ps, Gal_Shark_Diet_BT_PGCBLSGR.ps)
```

Remove all samples with no results (blacktips from PG, CB, LS, and GR)
```{r}
samples_to_remove <- c("con_G15", "con_G31", "con_G43", "G105", "G106", "G109", "G116", "G117", "G121", "G125", "G127", "G128", "G129", "G134", "G163", "G166", "G190", "G082", "G083", "G084", "G086", "G088", "G090", "G092", "G093", "G095", "G096", "G097", "G143", "G151", "G153", "G154", "G156", "G158", "G168", "G173", "G177", "G178", "G180", "G182", "G184", "G185")

Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps <- prune_samples(!(sample_names(Combined_Gal_Shark_Diet_PGCBLSGR.ps) %in% samples_to_remove), Combined_Gal_Shark_Diet_PGCBLSGR.ps)
```

Make files to upload to FigShare for journal submission
```{r}
Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.df <- as.data.frame(cbind(sample_data(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps)))
View(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.df)
write.csv(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.df, "./Chapter4sd")

Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.tt <- as.data.frame(cbind(tax_table(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps)))
View(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.tt)
write.csv(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.tt, "./Chapter4tt")

Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.otu <- as.data.frame(cbind(otu_table(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps)))
View(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.otu)
write.csv(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.otu, "./Chapter4otu")
```

Make a Final Taxa Table with the number of reads assigned to each genus or species (Blacktips)
```{r}
#merge motus by prey species - I don't think I need this because I did this in script 10
#NC_Shark_prey_species_merge.ps <- tax_glom(NC_Sharks1.ps, taxrank = "species", NArm = FALSE)

sample_variables(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps)
rank_names(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps)

#export read table with species or genus (if only genus available) instead of motu/asv now that they've been properly merged
taxtbl <- as.data.frame(tax_table(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps))
otutbl<-as.data.frame(otu_table(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps))
sampletbl<-as.data.frame(sample_data(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps))

otutbl$species <-
  taxtbl$species[match(rownames(otutbl),rownames(taxtbl))]

otutbl$genus <-
  taxtbl$genus[match(rownames(otutbl),rownames(taxtbl))]

df<-otutbl %>% relocate(genus)
df<-df %>% relocate(species, .after = genus)

df1<- as.data.frame(t(df))

df1<- rownames_to_column(df1, var="sample")
samples.df.1<-rownames_to_column(sampletbl, var="sample")
df2<- left_join(df1, samples.df.1, by ="sample")

colnames(df2)

df3<- df2 %>% select(-"type",-"control_type",-"nb_reads.x",-"nb_motus.x",-"seqdepth_ok.x",-"nb_reads_postmetabaR.x",-"nb_motus_postmetabaR.x",-"nb_reads.y",-"nb_motus.y",-"seqdepth_ok.y",-"nb_reads_postmetabaR.y",-"nb_motus_postmetabaR.y")

colnames(df3)

write.csv(df3, "Gal_BT_samples_by_species_combined_table.csv")

#Flip axis of the data frame
transformed_df3 <- t(df3)
write.csv(transformed_df3, "Gal_BT_samples_by_species_combined_table_flipped.csv")

################################################################################################################

#Make table that includes family and order too
otutbl$species <-
  taxtbl$species[match(rownames(otutbl),rownames(taxtbl))]

otutbl$genus <-
  taxtbl$genus[match(rownames(otutbl),rownames(taxtbl))]

otutbl$family <-
  taxtbl$family[match(rownames(otutbl),rownames(taxtbl))]

otutbl$order <-
  taxtbl$order[match(rownames(otutbl),rownames(taxtbl))]

df<-otutbl %>% relocate(species) %>% relocate(genus) %>% relocate(family) %>% relocate(order)
#df<-df %>% relocate(species, .after = genus)

df1<- as.data.frame(t(df))

df1<- rownames_to_column(df1, var="sample")
samples.df.1<-rownames_to_column(sampletbl, var="sample")
df2<- left_join(df1, samples.df.1, by ="sample")

colnames(df2)

df3<- df2 %>% select(-"type",-"control_type",-"nb_reads.x",-"nb_motus.x",-"seqdepth_ok.x",-"nb_reads_postmetabaR.x",-"nb_motus_postmetabaR.x",-"nb_reads.y",-"nb_motus.y",-"seqdepth_ok.y",-"nb_reads_postmetabaR.y",-"nb_motus_postmetabaR.y")

colnames(df3)

write.csv(df3, "Gal_BT_samples_by_species_combined_table_with familyand order.csv")

#edited_df3 <- read.csv("../Chapter3/Gal_BT_samples_by_species_combined_table_with familyand order_edited.csv")
#View(edited_df3)

#Flip axis of the data frame
#transformed_edited_df3 <- t(edited_df3)
#write.csv(transformed_edited_df3, "Gal_BT_samples_by_species_combined_table_with familyand order_edited_flipped.csv")
```

POO percentages for just scalloped hammerheads from PG and CB
```{r}
#Now only scalloped hammerheads - this is FOO
Combined_Gal_Shark_Diet.ps.SH<-subset_samples(Combined_Gal_Shark_Diet.ps,Species..scientific.name.=="Sphyrna lewini")

#Convert to presence absence: (FOO)
Combined_Gal_Shark_Diet.ps.SH.PA <- phyloseq_standardize_otu_abundance(Combined_Gal_Shark_Diet.ps.SH, method = "pa")

# Prune taxa with zero abundance across all samples
Combined_Gal_Shark_Diet.ps.SH.PA <- prune_taxa(
  taxa_sums(Combined_Gal_Shark_Diet.ps.SH.PA) > 0, 
  Combined_Gal_Shark_Diet.ps.SH.PA)

#Put it into dataframes
otu_df<-cbind(data.frame(otu_table(Combined_Gal_Shark_Diet.ps.SH.PA)))
otu_df$otu<-rownames(otu_df)

otu_df_long<-otu_df%>%
  pivot_longer(cols = -otu,names_to = "sample",values_to = "presence")

otu_df_long$sample <-gsub(pattern = ".",
                          "-",
                          x=otu_df_long$sample,
                          fixed = TRUE)

sample_df<-cbind(data.frame(sample_data(Combined_Gal_Shark_Diet.ps.SH.PA)))
sample_df<-rownames_to_column(sample_df, var = "sample")
sample_site_key<-sample_df%>%
  select(sample,Species..scientific.name.)

taxa_df<-cbind(data.frame(tax_table(Combined_Gal_Shark_Diet.ps.SH.PA)))
taxa_df<-rownames_to_column(taxa_df,var="otu")
otu_species_key<-taxa_df%>%
  select(otu,family,genus,species)

otu_df_long<-otu_df_long%>%
  left_join(sample_site_key)%>%
  left_join(otu_species_key)

#group by different taxonomic levels (shark species)

merge_shark.species_otu_counts<-otu_df_long%>%
  group_by(Species..scientific.name.,species)%>%
  summarise(total_count=sum(presence))%>%
  ungroup()

merge_shark.species_counts<-merge_shark.species_otu_counts%>%
  group_by(Species..scientific.name.)%>%
  mutate(total_site=sum(total_count))%>%
  ungroup()%>%
  mutate(rel_prct=(total_count/total_site)*100)

View(merge_shark.species_counts)
```

POO percentages for just blacktips from PG and CB
```{r}
#Now only blacktips - this is FOO
Combined_Gal_Shark_Diet.ps.BT<-subset_samples(Combined_Gal_Shark_Diet.ps,Species..scientific.name.=="Carcharhinus limbatus")

#Convert to presence absence: (FOO)
Combined_Gal_Shark_Diet.ps.BT.PA <- phyloseq_standardize_otu_abundance(Combined_Gal_Shark_Diet.ps.BT, method = "pa")

# Prune taxa with zero abundance across all samples
Combined_Gal_Shark_Diet.ps.BT.PA <- prune_taxa(
  taxa_sums(Combined_Gal_Shark_Diet.ps.BT.PA) > 0, 
  Combined_Gal_Shark_Diet.ps.BT.PA)

#Put it into dataframes
otu_df<-cbind(data.frame(otu_table(Combined_Gal_Shark_Diet.ps.BT.PA)))
otu_df$otu<-rownames(otu_df)

otu_df_long<-otu_df%>%
  pivot_longer(cols = -otu,names_to = "sample",values_to = "presence")

otu_df_long$sample <-gsub(pattern = ".",
                          "-",
                          x=otu_df_long$sample,
                          fixed = TRUE)

sample_df<-cbind(data.frame(sample_data(Combined_Gal_Shark_Diet.ps.BT.PA)))
sample_df<-rownames_to_column(sample_df, var = "sample")
sample_site_key<-sample_df%>%
  select(sample,Species..scientific.name.)

taxa_df<-cbind(data.frame(tax_table(Combined_Gal_Shark_Diet.ps.BT.PA)))
taxa_df<-rownames_to_column(taxa_df,var="otu")
otu_species_key<-taxa_df%>%
  select(otu,family,genus,species)

otu_df_long<-otu_df_long%>%
  left_join(sample_site_key)%>%
  left_join(otu_species_key)

#group by different taxonomic levels (shark species)

merge_shark.species_otu_counts<-otu_df_long%>%
  group_by(Species..scientific.name.,species)%>%
  summarise(total_count=sum(presence))%>%
  ungroup()

merge_shark.species_counts<-merge_shark.species_otu_counts%>%
  group_by(Species..scientific.name.)%>%
  mutate(total_site=sum(total_count))%>%
  ungroup()%>%
  mutate(rel_prct=(total_count/total_site)*100)

View(merge_shark.species_counts)
```

POO percentages for just blacktips from PG, LS, CB, GR
```{r}
#Now only blacktips - this is FOO
Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.BT<-subset_samples(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps,Species..scientific.name.=="Carcharhinus limbatus")

#Convert to presence absence: (FOO)
Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.BT.PA <- phyloseq_standardize_otu_abundance(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.BT, method = "pa")

# Prune taxa with zero abundance across all samples
Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.BT.PA <- prune_taxa(
  taxa_sums(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.BT.PA) > 0, 
  Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.BT.PA)

#Put it into dataframes
otu_df<-cbind(data.frame(otu_table(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.BT.PA)))
otu_df$otu<-rownames(otu_df)

otu_df_long<-otu_df%>%
  pivot_longer(cols = -otu,names_to = "sample",values_to = "presence")

otu_df_long$sample <-gsub(pattern = ".",
                          "-",
                          x=otu_df_long$sample,
                          fixed = TRUE)

sample_df<-cbind(data.frame(sample_data(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.BT.PA)))
sample_df<-rownames_to_column(sample_df, var = "sample")
sample_site_key<-sample_df%>%
  select(sample,Species..scientific.name.)

taxa_df<-cbind(data.frame(tax_table(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps.BT.PA)))
taxa_df<-rownames_to_column(taxa_df,var="otu")
otu_species_key<-taxa_df%>%
  select(otu,family,genus,species)

otu_df_long<-otu_df_long%>%
  left_join(sample_site_key)%>%
  left_join(otu_species_key)

#group by different taxonomic levels (shark species)

merge_shark.species_otu_counts<-otu_df_long%>%
  group_by(Species..scientific.name.,species)%>%
  summarise(total_count=sum(presence))%>%
  ungroup()

merge_shark.species_counts<-merge_shark.species_otu_counts%>%
  group_by(Species..scientific.name.)%>%
  mutate(total_site=sum(total_count))%>%
  ungroup()%>%
  mutate(rel_prct=(total_count/total_site)*100)

View(merge_shark.species_counts)
```

POO percentages with all sharks combined
```{r}
#Convert to presence absence: (FOO)
Combined_Gal_Shark_Diet.ps.PA <- phyloseq_standardize_otu_abundance(Combined_Gal_Shark_Diet.ps, method = "pa")

#### Puerto Grande ####

#Now only by site (Puerto Grande) - this is FOO
Combined_Gal_Shark_Diet.ps.PA.PuertoGrande<-subset_samples(Combined_Gal_Shark_Diet.ps.PA,Site.Name=="Puerto Grande")

# Prune taxa with zero abundance across all samples
Combined_Gal_Shark_Diet.ps.PA.PuertoGrande <- prune_taxa(
  taxa_sums(Combined_Gal_Shark_Diet.ps.PA.PuertoGrande) > 0, 
  Combined_Gal_Shark_Diet.ps.PA.PuertoGrande)

#Put it into dataframes
otu_df<-cbind(data.frame(otu_table(Combined_Gal_Shark_Diet.ps.PA.PuertoGrande)))
otu_df$otu<-rownames(otu_df)

otu_df_long<-otu_df%>%
  pivot_longer(cols = -otu,names_to = "sample",values_to = "presence")

otu_df_long$sample <-gsub(pattern = ".",
                          "-",
                          x=otu_df_long$sample,
                          fixed = TRUE)

sample_df<-cbind(data.frame(sample_data(Combined_Gal_Shark_Diet.ps.PA.PuertoGrande)))
sample_df<-rownames_to_column(sample_df, var = "sample")
sample_site_key<-sample_df%>%
  select(sample,Species..scientific.name.)

taxa_df<-cbind(data.frame(tax_table(Combined_Gal_Shark_Diet.ps.PA.PuertoGrande)))
taxa_df<-rownames_to_column(taxa_df,var="otu")
otu_species_key<-taxa_df%>%
  select(otu,family,genus,species)

otu_df_long<-otu_df_long%>%
  left_join(sample_site_key)%>%
  left_join(otu_species_key)

#group by different taxonomic levels (shark species)

merge_shark.species_otu_counts<-otu_df_long%>%
  group_by(Species..scientific.name.,species)%>%
  summarise(total_count=sum(presence))%>%
  ungroup()

merge_shark.species_counts<-merge_shark.species_otu_counts%>%
  group_by(Species..scientific.name.)%>%
  mutate(total_site=sum(total_count))%>%
  ungroup()%>%
  mutate(rel_prct=(total_count/total_site)*100)

View(merge_shark.species_counts)

#Custom order the prey species (separate custaceans and fishes)
# merge_shark.species_counts$species <- factor(
#   merge_shark.species_counts$species,
#   levels = c(
#     "Achelous tuberculatus", "Calappa convexa", "Meiosquilla dawsoni",
#     "Microcassiope xantusii", "Palaemon ritteri", "Abudefduf genus",
#     "Actinopteri class", "Carcharhinus genus", "Cyprinodon variegatus",
#     "Decapterus genus", "Lutjanus genus", "Mugil cephalus", "Mugil thoburni",
#     "Muraena lentiginosa", "Sphoeroides annulatus",
#     "Stegastes arcifrons", "Stegastes beebei"))

#plot 
POO_PG_plot <- ggplot(merge_shark.species_counts, aes(x = Species..scientific.name., y = rel_prct, fill = species)) +
  geom_bar(stat = "identity", position = "stack", color = NA) +  # Remove borders
  ylab("Relative Percent of Occurrence") +
  xlab("Shark Species") +
  theme_classic() +
  theme(text = element_text(size = 11)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(
    values = c(
      "Abudefduf genus" = "lightblue", "Achelous tuberculatus" = "tan",
      "Actinopteri class" = "navyblue", "Amphipholis squamata" = "hotpink1", 
      "Anisotremus scapularis" = "palegreen", "Calappa convexa" = "magenta4", "Caranx crysos" = "royalblue", 
      "Carcharhinus genus" = "grey", "Decapterus genus" = "indianred4", 
      "Euthynnus affinis" = "slategray2", "Fistularia commersonii" = "khaki", 
      "Grapsus grapsus" = "tomato", "Hemiramphus lutkei" = "green3", 
      "Lutjanus genus" = "aquamarine4", "Meiosquilla dawsoni" = "mediumpurple4", "Microcassiope xantusii" = "darkorange1", 
      "Mugil cephalus" = "blue", "Mugil thoburni" = "purple", "Muraena lentiginosa" = "lightpink", 
      "Pachygrapsus socius" = "goldenrod3", "Palaemon ritteri" = "coral3", "Sardinops genus" = "red", 
      "Scarus ghobban" = "yellow", "Selar crumenophthalmus" = "slateblue1", 
      "Semicossyphus pulcher" = "lawngreen", "Sphoeroides annulatus" = "lightseagreen", 
      "Sphyrna lewini" = "yellowgreen", "Stegastes arcifrons" = "darkgreen", 
      "Stegastes beebei" = "orange", "Teleophrys cristulipes" = "turquoise2", 
      "Thunnus genus" = "darksalmon"),
    labels = c(
      "Abudefduf genus" = "Abudefduf spp. (sergeant-majors)",
      "Achelous tuberculatus" = "Achelous tuberculatus (swimming crab)",
      "Actinopteri class" = "Actinopteri (ray-finned fish)",
      "Amphipholis squamata" = "Amphipholis squamata (dwarf brittle star)",
      "Anisotremus scapularis" = "Anisotremus scapularis (Peruvian grunt)",
      "Calappa convexa" = "Calappa convexa (shame faced crab)",
      "Caranx crysos" = "Caranx crysos (blue runner)",
      "Carcharhinus genus" = "Carcharhinus spp. (requiem sharks)",
      "Decapterus genus" = "Decapterus spp. (jack)",
      "Euthynnus affinis" = "Euthynnus affinis (mackerel tuna)",
      "Fistularia commersonii" = "Fistularia commersonii (bluespotted conetfish)",
      "Grapsus grapsus" = "Grapsus grapsus (Sally lightfoot crab)",
      "Hemiramphus lutkei" = "Hemiramphus lutkei (Lutke's halfbeak)",
      "Lutjanus genus" = "Lutjanus spp. (snapper)",
      "Meiosquilla dawsoni" = "Meiosquilla dawsoni (mantis shrimp)",
      "Microcassiope xantusii" = "Microcassiope xantusii (cassiope pebble crab)",
      "Mugil cephalus" = "Mugil cephalus (flathead grey mullet)",
      "Mugil thoburni" = "Mugil thoburni (Thoburn's mullet)",
      "Muraena lentiginosa" = "Muraena lentiginosa (jewel eel)",
      "Pachygrapsus socius" = "Pachygrapsus Socius (Pacific mottled shore crab)",
      "Palaemon ritteri" = "Palaemon ritteri (Barred grass shrimp)",
      "Sardinops genus" = "Sardinops spp. (sardine)",
      "Scarus ghobban" = "Scarus ghobban (blue-barred parrotfish)",
      "Selar crumenophthalmus" = "Selar crumenophthalmus (bigeye scad)",
      "Semicossyphus pulcher" = "Semicossyphus pulcher (California sheephead)",
      "Sphoeroides annulatus" = "Sphoeroides annulatus (bullseye puffer)",
      "Sphyrna lewini" = "Sphyrna lewini (scalloped hammerhead shark)",
      "Stegastes arcifrons" = "Stegastes arcifrons (yellowtail damselfish)",
      "Stegastes beebei" = "Stegastes beebei (Galapagos ringtail damselfish)",
      "Teleophrys cristulipes" = "Teleophrys cristulipes (spider crab)",
      "Thunnus genus" = "Thunnus spp. (tuna)"))

POO_PG_plot



#### Cartago Bay ####

#Now only by site (Cartago Bay) - this is FOO
Combined_Gal_Shark_Diet.ps.PA.CartagoBay<-subset_samples(Combined_Gal_Shark_Diet.ps.PA,Site.Name=="Cartago Bay")

# Prune taxa with zero abundance across all samples
Combined_Gal_Shark_Diet.ps.PA.CartagoBay <- prune_taxa(
  taxa_sums(Combined_Gal_Shark_Diet.ps.PA.CartagoBay) > 0, 
  Combined_Gal_Shark_Diet.ps.PA.CartagoBay)

#Put it into dataframes
otu_df<-cbind(data.frame(otu_table(Combined_Gal_Shark_Diet.ps.PA.CartagoBay)))
otu_df$otu<-rownames(otu_df)

otu_df_long<-otu_df%>%
  pivot_longer(cols = -otu,names_to = "sample",values_to = "presence")

otu_df_long$sample <-gsub(pattern = ".",
                          "-",
                          x=otu_df_long$sample,
                          fixed = TRUE)

sample_df<-cbind(data.frame(sample_data(Combined_Gal_Shark_Diet.ps.PA.CartagoBay)))
sample_df<-rownames_to_column(sample_df, var = "sample")
sample_site_key<-sample_df%>%
  select(sample,Species..scientific.name.)

taxa_df<-cbind(data.frame(tax_table(Combined_Gal_Shark_Diet.ps.PA.CartagoBay)))
taxa_df<-rownames_to_column(taxa_df,var="otu")
otu_species_key<-taxa_df%>%
  select(otu,family,genus,species)

otu_df_long<-otu_df_long%>%
  left_join(sample_site_key)%>%
  left_join(otu_species_key)

#group by different taxonomic levels (shark species)

merge_shark.species_otu_counts<-otu_df_long%>%
  group_by(Species..scientific.name.,species)%>%
  summarise(total_count=sum(presence))%>%
  ungroup()

merge_shark.species_counts<-merge_shark.species_otu_counts%>%
  group_by(Species..scientific.name.)%>%
  mutate(total_site=sum(total_count))%>%
  ungroup()%>%
  mutate(rel_prct=(total_count/total_site)*100)

View(merge_shark.species_counts)

#Custom order the prey species (separate custaceans and fishes)
# merge_shark.species_counts$species <- factor(
#   merge_shark.species_counts$species,
#   levels = c(
#     "Achelous tuberculatus", "Calappa convexa", "Meiosquilla dawsoni",
#     "Microcassiope xantusii", "Palaemon ritteri", "Abudefduf genus",
#     "Actinopteri class", "Carcharhinus genus", "Cyprinodon variegatus",
#     "Decapterus genus", "Lutjanus genus", "Mugil cephalus", "Mugil thoburni",
#     "Muraena lentiginosa", "Sphoeroides annulatus",
#     "Stegastes arcifrons", "Stegastes beebei"))

#plot 
POO_CB_plot <- ggplot(merge_shark.species_counts, aes(x = Species..scientific.name., y = rel_prct, fill = species)) +
  geom_bar(stat = "identity", position = "stack", color = NA) +  # Remove borders
  ylab("Relative Percent of Occurrence") +
  xlab("Shark Species") +
  theme_classic() +
  theme(text = element_text(size = 11)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(
    values = c(
      "Abudefduf genus" = "lightblue", "Achelous tuberculatus" = "tan",
      "Actinopteri class" = "navyblue", "Amphipholis squamata" = "hotpink1", 
      "Anisotremus scapularis" = "palegreen", "Calappa convexa" = "magenta4", "Caranx crysos" = "royalblue", 
      "Carcharhinus genus" = "grey", "Decapterus genus" = "indianred4", 
      "Euthynnus affinis" = "slategray2", "Fistularia commersonii" = "khaki", 
      "Grapsus grapsus" = "tomato", "Hemiramphus lutkei" = "green3", 
      "Lutjanus genus" = "aquamarine4", "Meiosquilla dawsoni" = "mediumpurple4", "Microcassiope xantusii" = "darkorange1", 
      "Mugil cephalus" = "blue", "Mugil thoburni" = "purple", "Muraena lentiginosa" = "lightpink", 
      "Pachygrapsus socius" = "goldenrod3", "Palaemon ritteri" = "coral3", "Sardinops genus" = "red", 
      "Scarus ghobban" = "yellow", "Selar crumenophthalmus" = "slateblue1", 
      "Semicossyphus pulcher" = "lawngreen", "Sphoeroides annulatus" = "lightseagreen", 
      "Sphyrna lewini" = "yellowgreen", "Stegastes arcifrons" = "darkgreen", 
      "Stegastes beebei" = "orange", "Teleophrys cristulipes" = "turquoise2", 
      "Thunnus genus" = "darksalmon"),
    labels = c(
      "Abudefduf genus" = "Abudefduf spp. (sergeant-majors)", 
      "Achelous tuberculatus" = "Achelous tuberculatus (swimming crab)",
      "Actinopteri class" = "Actinopteri (ray-finned fish)",
      "Amphipholis squamata" = "Amphipholis squamata (dwarf brittle star)",
      "Anisotremus scapularis" = "Anisotremus scapularis (Peruvian grunt)",
      "Calappa convexa" = "Calappa convexa (shame faced crab)",
      "Caranx crysos" = "Caranx crysos (blue runner)",
      "Carcharhinus genus" = "Carcharhinus spp. (requiem sharks)",
      "Decapterus genus" = "Decapterus spp. (jack)",
      "Euthynnus affinis" = "Euthynnus affinis (mackerel tuna)",
      "Fistularia commersonii" = "Fistularia commersonii (bluespotted conetfish)",
      "Grapsus grapsus" = "Grapsus grapsus (Sally lightfoot crab)",
      "Hemiramphus lutkei" = "Hemiramphus lutkei (Lutke's halfbeak)",
      "Lutjanus genus" = "Lutjanus spp. (snapper)",
      "Meiosquilla dawsoni" = "Meiosquilla dawsoni (mantis shrimp)",
      "Microcassiope xantusii" = "Microcassiope xantusii (cassiope pebble crab)",
      "Mugil cephalus" = "Mugil cephalus (flathead grey mullet)",
      "Mugil thoburni" = "Mugil thoburni (Thoburn's mullet)",
      "Muraena lentiginosa" = "Muraena lentiginosa (jewel eel)",
      "Pachygrapsus socius" = "Pachygrapsus Socius (Pacific mottled shore crab)",
      "Palaemon ritteri" = "Palaemon ritteri (Barred grass shrimp)",
      "Sardinops genus" = "Sardinops spp. (sardine)",
      "Scarus ghobban" = "Scarus ghobban (blue-barred parrotfish)",
      "Selar crumenophthalmus" = "Selar crumenophthalmus (bigeye scad)",
      "Semicossyphus pulcher" = "Semicossyphus pulcher (California sheephead)",
      "Sphoeroides annulatus" = "Sphoeroides annulatus (bullseye puffer)",
      "Sphyrna lewini" = "Sphyrna lewini (scalloped hammerhead shark)",
      "Stegastes arcifrons" = "Stegastes arcifrons (yellowtail damselfish)",
      "Stegastes beebei" = "Stegastes beebei (Galapagos ringtail damselfish)",
      "Teleophrys cristulipes" = "Teleophrys cristulipes (spider crab)",
      "Thunnus genus" = "Thunnus spp. (tuna)"))

POO_CB_plot

#########################################################################################

#Plot all together
library(gridExtra)
grid.arrange(POO_PG_plot, POO_CB_plot, ncol=1)
library(gridExtra)

grid.arrange(POO_PG_plot, POO_CB_plot, ncol = 1, heights = c(1, 1))


library(ggplot2)
library(cowplot)

# Extract the legend from one of the plots
legend <- get_legend(POO_PG_plot + theme(legend.position = "bottom"))

# Remove the legends from both plots
POO_PG_plot_no_legend <- POO_PG_plot + theme(legend.position = "none")
POO_CB_plot_no_legend <- POO_CB_plot + theme(legend.position = "none")

# Arrange the plots without the legend
combined_plots <- plot_grid(POO_PG_plot_no_legend, POO_CB_plot_no_legend, 
                            ncol = 1, align = "v", rel_heights = c(1, 1))

# Add the legend below the combined plots
final_plot <- plot_grid(combined_plots, legend, ncol = 1, rel_heights = c(1, 0.15))  # Adjust the 0.15 value for legend size

# Display the final plot with a single shared legend
final_plot
```

POO percentages with individual sharks (BC and PG)
```{r}
#Convert to presence absence: (FOO)
Combined_Gal_Shark_Diet.ps.PA <- phyloseq_standardize_otu_abundance(Combined_Gal_Shark_Diet_filtered.ps, method = "pa")

#### Scalloped hammerheads ####

#Now only by shark species (Scalloped Hammerheads) - this is FOO
Combined_Gal_Shark_Diet.ps.PA.SH<-subset_samples(Combined_Gal_Shark_Diet.ps.PA,Species..scientific.name.=="Sphyrna lewini")

# Prune taxa with zero abundance across all samples
Combined_Gal_Shark_Diet.ps.PA.SH <- prune_taxa(
  taxa_sums(Combined_Gal_Shark_Diet.ps.PA.SH) > 0, 
  Combined_Gal_Shark_Diet.ps.PA.SH)

#Put it into dataframes
otu_df<-cbind(data.frame(otu_table(Combined_Gal_Shark_Diet.ps.PA.SH)))
otu_df$otu<-rownames(otu_df)

otu_df_long<-otu_df%>%
  pivot_longer(cols = -otu,names_to = "sample",values_to = "presence")

otu_df_long$sample <-gsub(pattern = ".",
                          "-",
                          x=otu_df_long$sample,
                          fixed = TRUE)

sample_df<-cbind(data.frame(sample_data(Combined_Gal_Shark_Diet.ps.PA.SH)))
sample_df<-rownames_to_column(sample_df, var = "sample")
sample_site_key<-sample_df%>%
  select(sample,sample_id)

taxa_df<-cbind(data.frame(tax_table(Combined_Gal_Shark_Diet.ps.PA.SH)))
taxa_df<-rownames_to_column(taxa_df,var="otu")
otu_species_key<-taxa_df%>%
  select(otu,family,genus,species)

otu_df_long<-otu_df_long%>%
  left_join(sample_site_key)%>%
  left_join(otu_species_key)

#group by different taxonomic levels (shark species)

merge_shark.species_otu_counts<-otu_df_long%>%
  group_by(sample_id,species)%>%
  summarise(total_count=sum(presence))%>%
  ungroup()

merge_shark.species_counts<-merge_shark.species_otu_counts%>%
  group_by(sample_id)%>%
  mutate(total_site=sum(total_count))%>%
  ungroup()%>%
  mutate(rel_prct=(total_count/total_site)*100)

View(merge_shark.species_counts)

#Custom order the prey species (separate custaceans and fishes)
# merge_shark.species_counts$species <- factor(
#   merge_shark.species_counts$species,
#   levels = c(
#     "Achelous tuberculatus", "Calappa convexa", "Meiosquilla dawsoni",
#     "Microcassiope xantusii", "Palaemon ritteri", "Abudefduf genus",
#     "Actinopteri class", "Carcharhinus genus", "Cyprinodon variegatus",
#     "Decapterus genus", "Lutjanus genus", "Mugil cephalus", "Mugil thoburni",
#     "Muraena lentiginosa", "Sphoeroides annulatus",
#     "Stegastes arcifrons", "Stegastes beebei"))

#plot 
POO_SH_plot <- ggplot(merge_shark.species_counts, aes(x = sample_id, y = rel_prct, fill = species)) +
  geom_bar(stat = "identity", position = "stack", color = NA) +  # Remove borders
  ylab("Relative Percent of Occurrence") +
  xlab("Shark ID") +
  theme_classic() +
  theme(text = element_text(size = 11)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(
    values = c(
      "Abudefduf genus" = "lightblue", "Achelous tuberculatus" = "tan",
      "Actinopteri class" = "navyblue", "Amphipholis squamata" = "hotpink1", 
      "Anisotremus scapularis" = "palegreen", "Calappa convexa" = "magenta4", "Caranx crysos" = "royalblue", 
      "Carcharhinus genus" = "grey", "Decapterus genus" = "indianred4", 
      "Euthynnus affinis" = "slategray2", "Fistularia commersonii" = "khaki", 
      "Grapsus grapsus" = "tomato", "Hemiramphus lutkei" = "green3", 
      "Lutjanus genus" = "aquamarine4", "Meiosquilla dawsoni" = "mediumpurple4", "Microcassiope xantusii" = "darkorange1", 
      "Mugil cephalus" = "blue", "Mugil thoburni" = "purple", "Muraena lentiginosa" = "lightpink", 
      "Pachygrapsus socius" = "goldenrod3", "Palaemon ritteri" = "coral3", "Sardinops genus" = "red", 
      "Scarus ghobban" = "yellow", "Selar crumenophthalmus" = "slateblue1", 
      "Semicossyphus pulcher" = "lawngreen", "Sphoeroides annulatus" = "lightseagreen", 
      "Sphyrna lewini" = "yellowgreen", "Stegastes arcifrons" = "darkgreen", 
      "Stegastes beebei" = "orange", "Teleophrys cristulipes" = "turquoise2", 
      "Thunnus genus" = "darksalmon"),
    labels = c(
      "Abudefduf genus" = "Abudefduf spp. (sergeant-majors)", 
      "Achelous tuberculatus" = "Achelous tuberculatus (swimming crab)",
      "Actinopteri class" = "Actinopteri (ray-finned fish)",
      "Amphipholis squamata" = "Amphipholis squamata (dwarf brittle star)",
      "Anisotremus scapularis" = "Anisotremus scapularis (Peruvian grunt)",
      "Calappa convexa" = "Calappa convexa (shame faced crab)",
      "Caranx crysos" = "Caranx crysos (blue runner)",
      "Carcharhinus genus" = "Carcharhinus spp. (requiem sharks)",
      "Decapterus genus" = "Decapterus spp. (jack)",
      "Euthynnus affinis" = "Euthynnus affinis (mackerel tuna)",
      "Fistularia commersonii" = "Fistularia commersonii (bluespotted conetfish)",
      "Grapsus grapsus" = "Grapsus grapsus (Sally lightfoot crab)",
      "Hemiramphus lutkei" = "Hemiramphus lutkei (Lutke's halfbeak)",
      "Lutjanus genus" = "Lutjanus spp. (snapper)",
      "Meiosquilla dawsoni" = "Meiosquilla dawsoni (mantis shrimp)",
      "Microcassiope xantusii" = "Microcassiope xantusii (cassiope pebble crab)",
      "Mugil cephalus" = "Mugil cephalus (flathead grey mullet)",
      "Mugil thoburni" = "Mugil thoburni (Thoburn's mullet)",
      "Muraena lentiginosa" = "Muraena lentiginosa (jewel eel)",
      "Pachygrapsus socius" = "Pachygrapsus Socius (Pacific mottled shore crab)",
      "Palaemon ritteri" = "Palaemon ritteri (Barred grass shrimp)",
      "Sardinops genus" = "Sardinops spp. (sardine)",
      "Scarus ghobban" = "Scarus ghobban (blue-barred parrotfish)",
      "Selar crumenophthalmus" = "Selar crumenophthalmus (bigeye scad)",
      "Semicossyphus pulcher" = "Semicossyphus pulcher (California sheephead)",
      "Sphoeroides annulatus" = "Sphoeroides annulatus (bullseye puffer)",
      "Sphyrna lewini" = "Sphyrna lewini (scalloped hammerhead shark)",
      "Stegastes arcifrons" = "Stegastes arcifrons (yellowtail damselfish)",
      "Stegastes beebei" = "Stegastes beebei (Galapagos ringtail damselfish)",
      "Teleophrys cristulipes" = "Teleophrys cristulipes (spider crab)",
      "Thunnus genus" = "Thunnus spp. (tuna)"))

POO_SH_plot



#### Cartago Bay ####

#Now only by shark species (Blacktip) - this is FOO
Combined_Gal_Shark_Diet.ps.PA.BT<-subset_samples(Combined_Gal_Shark_Diet_filtered.ps,Species..scientific.name.=="Carcharhinus limbatus")

# Prune taxa with zero abundance across all samples
Combined_Gal_Shark_Diet.ps.PA.BT <- prune_taxa(
  taxa_sums(Combined_Gal_Shark_Diet.ps.PA.BT) > 0, 
  Combined_Gal_Shark_Diet.ps.PA.BT)

#Put it into dataframes
otu_df<-cbind(data.frame(otu_table(Combined_Gal_Shark_Diet.ps.PA.BT)))
otu_df$otu<-rownames(otu_df)

otu_df_long<-otu_df%>%
  pivot_longer(cols = -otu,names_to = "sample",values_to = "presence")

otu_df_long$sample <-gsub(pattern = ".",
                          "-",
                          x=otu_df_long$sample,
                          fixed = TRUE)

sample_df<-cbind(data.frame(sample_data(Combined_Gal_Shark_Diet.ps.PA.BT)))
sample_df<-rownames_to_column(sample_df, var = "sample")
sample_site_key<-sample_df%>%
  select(sample,sample_id)

taxa_df<-cbind(data.frame(tax_table(Combined_Gal_Shark_Diet.ps.PA.BT)))
taxa_df<-rownames_to_column(taxa_df,var="otu")
otu_species_key<-taxa_df%>%
  select(otu,family,genus,species)

otu_df_long<-otu_df_long%>%
  left_join(sample_site_key)%>%
  left_join(otu_species_key)

#group by different taxonomic levels (shark species)

merge_shark.species_otu_counts<-otu_df_long%>%
  group_by(sample_id,species)%>%
  summarise(total_count=sum(presence))%>%
  ungroup()

merge_shark.species_counts<-merge_shark.species_otu_counts%>%
  group_by(sample_id)%>%
  mutate(total_site=sum(total_count))%>%
  ungroup()%>%
  mutate(rel_prct=(total_count/total_site)*100)

View(merge_shark.species_counts)

#Custom order the prey species (separate custaceans and fishes)
# merge_shark.species_counts$species <- factor(
#   merge_shark.species_counts$species,
#   levels = c(
#     "Achelous tuberculatus", "Calappa convexa", "Meiosquilla dawsoni",
#     "Microcassiope xantusii", "Palaemon ritteri", "Abudefduf genus",
#     "Actinopteri class", "Carcharhinus genus", "Cyprinodon variegatus",
#     "Decapterus genus", "Lutjanus genus", "Mugil cephalus", "Mugil thoburni",
#     "Muraena lentiginosa", "Sphoeroides annulatus",
#     "Stegastes arcifrons", "Stegastes beebei"))

#plot 
POO_BT_plot <- ggplot(merge_shark.species_counts, aes(x = sample_id, y = rel_prct, fill = species)) +
  geom_bar(stat = "identity", position = "stack", color = NA) +  # Remove borders
  ylab("Relative Percent of Occurrence") +
  xlab("Shark ID") +
  theme_classic() +
  theme(text = element_text(size = 11)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(
    values = c(
      "Abudefduf genus" = "lightblue", "Achelous tuberculatus" = "tan",
      "Actinopteri class" = "navyblue", "Amphipholis squamata" = "hotpink1", 
      "Anisotremus scapularis" = "palegreen", "Calappa convexa" = "magenta4", "Caranx crysos" = "royalblue", 
      "Carcharhinus genus" = "grey", "Decapterus genus" = "indianred4", 
      "Euthynnus affinis" = "slategray2", "Fistularia commersonii" = "khaki", 
      "Grapsus grapsus" = "tomato", "Hemiramphus lutkei" = "green3", 
      "Lutjanus genus" = "aquamarine4", "Meiosquilla dawsoni" = "mediumpurple4", "Microcassiope xantusii" = "darkorange1", 
      "Mugil cephalus" = "blue", "Mugil thoburni" = "purple", "Muraena lentiginosa" = "lightpink", 
      "Pachygrapsus socius" = "goldenrod3", "Palaemon ritteri" = "coral3", "Sardinops genus" = "red", 
      "Scarus ghobban" = "yellow", "Selar crumenophthalmus" = "slateblue1", 
      "Semicossyphus pulcher" = "lawngreen", "Sphoeroides annulatus" = "lightseagreen", 
      "Sphyrna lewini" = "yellowgreen", "Stegastes arcifrons" = "darkgreen", 
      "Stegastes beebei" = "orange", "Teleophrys cristulipes" = "turquoise2", 
      "Thunnus genus" = "darksalmon"),
    labels = c(
      "Abudefduf genus" = "Abudefduf spp. (sergeant-majors)", 
      "Achelous tuberculatus" = "Achelous tuberculatus (swimming crab)",
      "Actinopteri class" = "Actinopteri (ray-finned fish)",
      "Amphipholis squamata" = "Amphipholis squamata (dwarf brittle star)",
      "Anisotremus scapularis" = "Anisotremus scapularis (Peruvian grunt)",
      "Calappa convexa" = "Calappa convexa (shame faced crab)",
      "Caranx crysos" = "Caranx crysos (blue runner)",
      "Carcharhinus genus" = "Carcharhinus spp. (requiem sharks)",
      "Decapterus genus" = "Decapterus spp. (jack)",
      "Euthynnus affinis" = "Euthynnus affinis (mackerel tuna)",
      "Fistularia commersonii" = "Fistularia commersonii (bluespotted conetfish)",
      "Grapsus grapsus" = "Grapsus grapsus (Sally lightfoot crab)",
      "Hemiramphus lutkei" = "Hemiramphus lutkei (Lutke's halfbeak)",
      "Lutjanus genus" = "Lutjanus spp. (snapper)",
      "Meiosquilla dawsoni" = "Meiosquilla dawsoni (mantis shrimp)",
      "Microcassiope xantusii" = "Microcassiope xantusii (cassiope pebble crab)",
      "Mugil cephalus" = "Mugil cephalus (flathead grey mullet)",
      "Mugil thoburni" = "Mugil thoburni (Thoburn's mullet)",
      "Muraena lentiginosa" = "Muraena lentiginosa (jewel eel)",
      "Pachygrapsus socius" = "Pachygrapsus Socius (Pacific mottled shore crab)",
      "Palaemon ritteri" = "Palaemon ritteri (Barred grass shrimp)",
      "Sardinops genus" = "Sardinops spp. (sardine)",
      "Scarus ghobban" = "Scarus ghobban (blue-barred parrotfish)",
      "Selar crumenophthalmus" = "Selar crumenophthalmus (bigeye scad)",
      "Semicossyphus pulcher" = "Semicossyphus pulcher (California sheephead)",
      "Sphoeroides annulatus" = "Sphoeroides annulatus (bullseye puffer)",
      "Sphyrna lewini" = "Sphyrna lewini (scalloped hammerhead shark)",
      "Stegastes arcifrons" = "Stegastes arcifrons (yellowtail damselfish)",
      "Stegastes beebei" = "Stegastes beebei (Galapagos ringtail damselfish)",
      "Teleophrys cristulipes" = "Teleophrys cristulipes (spider crab)",
      "Thunnus genus" = "Thunnus spp. (tuna)"))

POO_BT_plot

#########################################################################################

#Plot all together
library(gridExtra)
grid.arrange(POO_SH_plot, POO_BT_plot, ncol=1)
library(gridExtra)

grid.arrange(POO_SH_plot, POO_BT_plot, ncol = 1, heights = c(1, 1))


library(ggplot2)
library(cowplot)

# Extract the legend from one of the plots
legend <- get_legend(POO_SH_plot + theme(legend.position = "bottom"))

# Remove the legends from both plots
POO_SH_plot_no_legend <- POO_SH_plot + theme(legend.position = "none")
POO_BT_plot_no_legend <- POO_BT_plot + theme(legend.position = "none")

# Arrange the plots without the legend
combined_plots <- plot_grid(POO_SH_plot_no_legend, POO_BT_plot_no_legend, 
                            ncol = 1, align = "v", rel_heights = c(1, 1))

# Add the legend below the combined plots
final_plot <- plot_grid(combined_plots, legend, ncol = 1, rel_heights = c(1, 0.15))  # Adjust the 0.15 value for legend size

# Display the final plot with a single shared legend
final_plot
```

MDS Plots (prey taxa composition between sties)
```{r}
# Removing the NAs (anything that wasn't IDed to genus gets deleted here)
Combined_Gal_Shark_Diet_filtered.ps_species.ps<-tax_glom(Combined_Gal_Shark_Diet_filtered.ps, taxrank = "species", NArm = TRUE)

# play with ordinations interactively
pseq_prey_species <- Combined_Gal_Shark_Diet_filtered.ps_species.ps %>%
  phyloseq_validate()

# MDS Plot (genus) - jsd, with circles
pseq_prey_species %>%
  tax_transform(rank = "species", trans = "compositional") %>%
  dist_calc(dist = "jsd") %>%
  ord_calc(method = "PCoA") %>% 
  ord_plot(
    axes = c(1, 2),
    colour = "Species..scientific.name.",
    fill = "Species..scientific.name.",
    shape = "Site.Name",
    alpha = 0.5,
    size = 2
  ) +
  ggplot2::stat_ellipse(
    ggplot2::aes(colour = Species..scientific.name.)
  ) +
  theme(
    panel.grid = element_blank(),
    panel.grid.major = element_blank())

# MDS Plot (genus) - jsd, with circles
pseq_prey_species %>%
  tax_transform(rank = "species", trans = "compositional") %>%
  dist_calc(dist = "jsd") %>%
  ord_calc(method = "PCoA") %>% 
  ord_plot(
    axes = c(1, 2),
    colour = "Site.Name",
    fill = "Site.Name",
    shape = "Species..scientific.name.",
    alpha = 0.5,
    size = 2
  ) +
  ggplot2::stat_ellipse(
    ggplot2::aes(colour = Species..scientific.name.)
  ) +
  theme(
    panel.grid = element_blank(),
    panel.grid.major = element_blank())

#PCA by site and shark species
# Define a custom color palette (adjust as needed)
my_colors <- c("Sphyrna lewini" = "yellowgreen", "Carcharhinus limbatus" = "royalblue")

# Modify your plot pipeline
pseq_prey_species %>%
  tax_transform(rank = "genus", trans = "compositional") %>%
  ord_calc(method = "auto") %>%
  ord_plot(
    axes = c(1, 2),
    plot_taxa = 1:5,
    colour = "Species..scientific.name.", 
    fill = "Species..scientific.name.",
    shape = "Site.Name", 
    alpha = 0.5,
    size = 2
  ) +
  scale_color_manual(values = my_colors) + 
  scale_fill_manual(values = my_colors)



##Stats - Between sites (no statistical difference in prey composition)
jsd_dist_matrix <- phyloseq::distance(pseq_prey_species, method ="jsd")

set.seed(200)
permanova <- vegan::adonis2(jsd_dist_matrix ~ phyloseq::sample_data(pseq_prey_species)$Species..scientific.name.)
permanova
View(permanova)
```

Calculate Levin's niche width and Levin's niche overlap - Shark Species 
```{r}
# Merge data by site name
Combined_Gal_Shark_Diet_filtered.ps_SharkSpecies <- merge_samples(Combined_Gal_Shark_Diet_filtered.ps, group = "Species..scientific.name.") 

# Ensure the OTU table has samples as rows and taxa as columns
Combined_Gal_Shark_Diet_filtered.ps_SharkSpecies_Reverse <- t(otu_table(Combined_Gal_Shark_Diet_filtered.ps_SharkSpecies, taxa_are_rows = TRUE))

# Convert to a data frame
Combined_Gal_Shark_Diet_filtered.ps_SharkSpecies_Reverse <- as.data.frame(Combined_Gal_Shark_Diet_filtered.ps_SharkSpecies_Reverse)

# View the structure to confirm samples are rows
View(Combined_Gal_Shark_Diet_filtered.ps_SharkSpecies_Reverse)

# Calculate niche width for samples
levins.niche.width_species <- niche.width(Combined_Gal_Shark_Diet_filtered.ps_SharkSpecies_Reverse, method = "levins")
View(levins.niche.width_species)

# Calculate niche overlap for samples
levins.niche.overlap_species <- niche.overlap(Combined_Gal_Shark_Diet_filtered.ps_SharkSpecies_Reverse, method = "levins")
print(levins.niche.overlap_species)
```

Calculate Levin's niche width - Scalloped hammerhead by site
```{r}
# Separate by shark species
Combined_Gal_Shark_Diet_filtered.ps_SH<-subset_samples(Combined_Gal_Shark_Diet_filtered.ps,Species..scientific.name.=="Sphyrna lewini")
View(Combined_Gal_Shark_Diet_filtered.ps_SH)

# Merge data by site name
Combined_Gal_Shark_Diet_filtered.ps_SH_stie <- merge_samples(Combined_Gal_Shark_Diet_filtered.ps_SH, group = "Site.Name")

# Calculate richness
richness_data <- estimate_richness(Combined_Gal_Shark_Diet_filtered.ps_SH_stie, measures = c("Observed"))

# Print richness per sample
print(richness_data)

# Ensure the OTU table has samples as rows and taxa as columns
Combined_Gal_Shark_Diet_filtered.ps_SH_stie_Reverse <- t(otu_table(Combined_Gal_Shark_Diet_filtered.ps_SH_stie, taxa_are_rows = TRUE))

# Convert to a data frame
Combined_Gal_Shark_Diet_filtered.ps_SH_stie_Reverse <- as.data.frame(Combined_Gal_Shark_Diet_filtered.ps_SH_stie_Reverse)

# View the structure to confirm samples are rows
View(Combined_Gal_Shark_Diet_filtered.ps_SH_stie_Reverse)

# Calculate niche width for samples
levins.niche.width_site <- niche.width(Combined_Gal_Shark_Diet_filtered.ps_SH_stie_Reverse, method = "levins")
View(levins.niche.width_site)

# Calculate niche overlap for samples
levins.niche.overlap_site_SH <- niche.overlap(Combined_Gal_Shark_Diet_filtered.ps_SH_stie_Reverse, method = "levins")
print(levins.niche.overlap_site_SH)
```

MDS Plots (prey taxa composition between sties) Scalloped hammerheads
```{r}
# Removing the NAs (anything that wasn't IDed to genus gets deleted here)
Gal_Shark_Diet_edited.ps_species.ps<-tax_glom(Combined_Gal_Shark_Diet_filtered.ps_SH, taxrank = "species", NArm = TRUE)

# play with ordinations interactively
pseq_SH_prey_species <- Gal_Shark_Diet_edited.ps_species.ps %>%
  phyloseq_validate()

# MDS Plot (genus) - jsd, with circles
pseq_SH_prey_species %>%
  tax_transform(rank = "species", trans = "compositional") %>%
  dist_calc(dist = "jsd") %>%
  ord_calc(method = "PCoA") %>% 
  ord_plot(
    axes = c(1, 2),
    colour = "Site.Name",
    fill = "Site.Name",
    shape = "circle",
    alpha = 0.5,
    size = 2
  ) +
  ggplot2::stat_ellipse(
    ggplot2::aes(colour = Site.Name)
  ) +
  theme(
    panel.grid = element_blank(),
    panel.grid.major = element_blank())


##Stats - Between sites (no statistical difference)
jsd_dist_matrix <- phyloseq::distance(pseq_SH_prey_species, method ="jsd")

set.seed(200)
permanova <- vegan::adonis2(jsd_dist_matrix ~ phyloseq::sample_data(pseq_SH_prey_species)$Site.Name)
permanova
View(permanova)
```

Calculate Levin's niche width - blacktips by site (only PG and CB)
```{r}
# Separate by shark species
Combined_Gal_Shark_Diet_filtered.ps_BT<-subset_samples(Combined_Gal_Shark_Diet_filtered.ps,Species..scientific.name.=="Carcharhinus limbatus")
View(Combined_Gal_Shark_Diet_filtered.ps_BT)

# Merge data by site name
Combined_Gal_Shark_Diet_filtered.ps_BT_stie <- merge_samples(Combined_Gal_Shark_Diet_filtered.ps_BT, group = "Site.Name")

# Calculate richness
richness_data <- estimate_richness(Combined_Gal_Shark_Diet_filtered.ps_BT_stie, measures = c("Observed"))

# Print richness per sample
print(richness_data)

# Ensure the OTU table has samples as rows and taxa as columns
Combined_Gal_Shark_Diet_filtered.ps_BT_stie_Reverse <- t(otu_table(Combined_Gal_Shark_Diet_filtered.ps_BT_stie, taxa_are_rows = TRUE))

# Convert to a data frame
Combined_Gal_Shark_Diet_filtered.ps_BT_stie_Reverse <- as.data.frame(Combined_Gal_Shark_Diet_filtered.ps_BT_stie_Reverse)

# View the structure to confirm samples are rows
View(Combined_Gal_Shark_Diet_filtered.ps_BT_stie_Reverse)

# Calculate niche width for samples
levins.niche.width_site_BT <- niche.width(Combined_Gal_Shark_Diet_filtered.ps_BT_stie_Reverse, method = "levins")
View(levins.niche.width_site_BT)

# Calculate niche overlap for samples
levins.niche.overlap_site_BT <- niche.overlap(Combined_Gal_Shark_Diet_filtered.ps_BT_stie_Reverse, method = "levins")
print(levins.niche.overlap_site_BT)
```

MDS Plots (prey taxa composition between sties) Blacktips
```{r}
# Removing the NAs (anything that wasn't IDed to genus gets deleted here)
Gal_Shark_Diet_edited.ps_species.ps<-tax_glom(Combined_Gal_Shark_Diet_filtered.ps_BT, taxrank = "species", NArm = TRUE)

# play with ordinations interactively
pseq_BT_prey_species <- Gal_Shark_Diet_edited.ps_species.ps %>%
  phyloseq_validate()

# MDS Plot (genus) - jsd, with circles
pseq_BT_prey_species %>%
  tax_transform(rank = "species", trans = "compositional") %>%
  dist_calc(dist = "jsd") %>%
  ord_calc(method = "PCoA") %>% 
  ord_plot(
    axes = c(1, 2),
    colour = "Site.Name",
    fill = "Site.Name",
    shape = "circle",
    alpha = 0.5,
    size = 2
  ) +
  ggplot2::stat_ellipse(
    ggplot2::aes(colour = Site.Name)
  ) +
  theme(
    panel.grid = element_blank(),
    panel.grid.major = element_blank())


##Stats - Between sites (no statistical difference)
jsd_dist_matrix <- phyloseq::distance(pseq_BT_prey_species, method ="jsd")

set.seed(200)
permanova <- vegan::adonis2(jsd_dist_matrix ~ phyloseq::sample_data(pseq_BT_prey_species)$Site.Name)
permanova
View(permanova)
```

MDS Plots (prey taxa composition per individual) Blacktips from PG and LS
```{r}
# Separate by shark species
Combined_Gal_Shark_Diet_PGLS_filtered.ps_BT<-subset_samples(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps, Island=="San Cristobal")
View(Combined_Gal_Shark_Diet_PGLS_filtered.ps_BT)

# Removing the NAs (anything that wasn't IDed to genus gets deleted here)
Combined_Gal_Shark_Diet_PGLS_filtered.ps_BT<-tax_glom(Combined_Gal_Shark_Diet_PGLS_filtered.ps_BT, taxrank = "species", NArm = TRUE)

# play with ordinations interactively
pseq_BT_prey_SanCristobal <- Combined_Gal_Shark_Diet_PGLS_filtered.ps_BT %>%
  phyloseq_validate()

# MDS Plot (genus) - jsd, with circles
pseq_BT_prey_SanCristobal %>%
  tax_transform(rank = "species", trans = "compositional") %>%
  dist_calc(dist = "jsd") %>%
  ord_calc(method = "PCoA") %>% 
  ord_plot(
    axes = c(1, 2),
    colour = "Site.Name",
    fill = "Site.Name",
    shape = "circle",
    alpha = 0.5,
    size = 2
  ) +
  ggplot2::stat_ellipse(
    ggplot2::aes(colour = Site.Name)
  ) +
  theme(
    panel.grid = element_blank(),
    panel.grid.major = element_blank())
```

Calculate Levin's niche width - blacktips by site
```{r}
# Separate by shark species
Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT<-subset_samples(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps,Species..scientific.name.=="Carcharhinus limbatus")
View(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT)

# Merge data by site name
Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_stie <- merge_samples(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT, group = "Site.Name")

# Calculate richness
richness_data <- estimate_richness(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_stie, measures = c("Observed"))

# Print richness per sample
print(richness_data)

# Ensure the OTU table has samples as rows and taxa as columns
Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_stie_Reverse <- t(otu_table(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_stie, taxa_are_rows = TRUE))

# Convert to a data frame
Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_stie_Reverse <- as.data.frame(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_stie_Reverse)

# View the structure to confirm samples are rows
View(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_stie_Reverse)

# Calculate niche width for samples
levins.niche.width_site_BT <- niche.width(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_stie_Reverse, method = "levins")
View(levins.niche.width_site_BT)
```

Calculate Levin's niche width - blacktips by individual to look at individual specialization
```{r}
# Separate by shark species
Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT<-subset_samples(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps,Species..scientific.name.=="Carcharhinus limbatus")
View(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT)

# Merge data by site name
Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_individual <- merge_samples(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT, group = "sample_id")

# Calculate richness
richness_data <- estimate_richness(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_individual, measures = c("Observed"))

# Print richness per sample
print(richness_data)

# Ensure the OTU table has samples as rows and taxa as columns
Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_individual_Reverse <- t(otu_table(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_individual, taxa_are_rows = TRUE))

# Convert to a data frame
Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_individual_Reverse <- as.data.frame(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_individual_Reverse)

# View the structure to confirm samples are rows
View(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_individual_Reverse)

# Calculate niche width for samples
levins.niche.width_individual_BT <- niche.width(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps_BT_individual_Reverse, method = "levins")
View(levins.niche.width_individual_BT)
```

Richness Boxplot (only sharded nurseries)
```{r}
library(RColorBrewer)
species.richness_shark.species <- plot_richness(
  Combined_Gal_Shark_Diet_filtered.ps,
  x = "Site.Name",
  color = "Species..scientific.name.",
  measures = c("Observed")
) +
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +  # Slightly transparent boxplot, hide outliers
  geom_jitter(width = 0.2, size = 2, alpha = 0.8) +  # Jitter points for visibility
  theme_classic() +
  theme(axis.text.x = element_text(angle = 0)) +
  scale_color_brewer(palette = "Set2")

species.richness_shark.species
```

Richness Boxplot (all four bays)
```{r}
library(RColorBrewer)

species.richness_all.sites <- plot_richness(
  Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps,
  x = "Site.Name",
  color = "Species..scientific.name.",
  measures = c("Observed")
) +
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +  # Slightly transparent boxplot, hide outliers
  geom_jitter(width = 0.2, size = 2, alpha = 0.8) +  # Jitter points for visibility
  theme_classic() +
  theme(axis.text.x = element_text(angle = 0)) +
  scale_color_brewer(palette = "Set2")

species.richness_all.sites


#####Now calculate with 95% CI
library(phyloseq)
library(dplyr)
library(ggplot2)

# Calculate observed richness from phyloseq object
richness_data <- estimate_richness(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps, measures = "Observed")

# Extract sample metadata (Site Name & Island info)
sample_metadata <- as.data.frame(sample_data(Combined_Gal_Shark_Diet_PGCBLSGR_filtered.ps))

# Merge richness data with metadata
richness_data$Site.Name <- sample_metadata$Site.Name
richness_data$Species..scientific.name. <- sample_metadata$Species..scientific.name.

# Compute summary statistics per site-shark species combo
richness_summary <- richness_data %>%
  group_by(Site.Name, Species..scientific.name.) %>%
  summarise(
    mean_richness = mean(Observed, na.rm = TRUE),
    sd_richness = sd(Observed, na.rm = TRUE),
    n = n(),
    se = sd_richness / sqrt(n),
    lower_CI = mean_richness - (1.96 * se),
    upper_CI = mean_richness + (1.96 * se),
    .groups = "drop"
  )

print(richness_summary)  # Check the calculated confidence intervals

# Define a custom order for the x-axis (keeping shark species within the same site together)
custom_x_order <- richness_data %>%
  arrange(Site.Name, Species..scientific.name.) %>%  # Sort by Site first, then species
  mutate(x_factor = factor(interaction(Site.Name, Species..scientific.name.), 
                           levels = unique(interaction(Site.Name, Species..scientific.name.))))  # Set factor levels

# Update the richness summary to match the new order
richness_summary <- richness_summary %>%
  mutate(x_factor = factor(interaction(Site.Name, Species..scientific.name.), 
                           levels = levels(custom_x_order$x_factor)))  # Ensure same ordering

# Create base ggplot with the ordered x-axis
species_richness_plot <- ggplot(custom_x_order, aes(
  x = x_factor,  # Use the ordered factor for x-axis
  y = Observed,
  color = Species..scientific.name.
)) +
  geom_jitter(size = 2, alpha = 0.8, width = 0.2) +  # Jitter points for individual sharks
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +  # Boxplot for spread
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_brewer(palette = "Set2") +
  scale_x_discrete(name = "Site & Shark Species") +  # Keep site-species labels
  labs(y = "Observed Richness")

# Add confidence interval error bars
species_richness_plot <- species_richness_plot +
  geom_errorbar(
    data = richness_summary,
    aes(
      x = x_factor, 
      ymin = lower_CI, 
      ymax = upper_CI, 
      color = Species..scientific.name.
    ),
    width = 0.2, inherit.aes = FALSE, color = "grey"
  ) +
  geom_point(  # Add mean points for clarity
    data = richness_summary,
    aes(
      x = x_factor, 
      y = mean_richness, 
      color = Species..scientific.name.
    ),
    size = 3, 
    shape = 18  # Triangle shape for means
  )

species_richness_plot
```

ANOVA comparing the richness between shark species and nursery bay
```{r}
# Calculate richness (Observed richness as an example)
alpha_div <- estimate_richness(Combined_Gal_Shark_Diet_filtered.ps, measures = "Observed")

# Convert sample metadata to a dataframe
metadata <- as.data.frame(sample_data(Combined_Gal_Shark_Diet_filtered.ps))

# Combine richness and metadata
alpha_data <- cbind(metadata, Observed = alpha_div$Observed)

# Run two-way ANOVA (Shark species & Site)
anova_result <- aov(Observed ~ Species..scientific.name. * Site.Name, data = alpha_data)

# View results
summary(anova_result)
```

Rarefaction curve - SH
```{r}
Rarefaction_SH <- read.csv("Rarefaction_Gal_SH.csv")
#View(Rarefaction_SH)

#Switch axis
Rarefaction_SH_Reverse <- data.frame(t(Rarefaction_SH[-1]))
colnames(Rarefaction_SH_Reverse) <- Rarefaction_SH[, 1]
View(Rarefaction_SH_Reverse)

sample_data(Rarefaction_SH_Reverse)
summary(Rarefaction_SH_Reverse)
attach(Rarefaction_SH_Reverse)
names(Rarefaction_SH_Reverse)

#ci=2 means that the CI is 95%
sp_SH <- specaccum(Rarefaction_SH_Reverse, object="prey taxa", permutations = 100, ci=2)
#View(sp_BT)

plot(sp_SH, ylab="Number of Prey Taxa", xlab="Number of samples",col="black",ci.type="poly", ci.col="grey", ci.lty = 0,
     xlim=c(1, 32), ylim=c(1, 20))
```

Rarefaction curve - BT (PG and CB)
```{r}
Rarefaction_BT_PGCB <- read.csv("Rarefaction_Gal_BT_PGCB.csv")
#View(Rarefaction_BT_PGCB)

#Switch axis
Rarefaction_BT_PGCB_Reverse <- data.frame(t(Rarefaction_BT_PGCB[-1]))
colnames(Rarefaction_BT_PGCB_Reverse) <- Rarefaction_BT_PGCB[, 1]
View(Rarefaction_BT_PGCB_Reverse)

sample_data(Rarefaction_BT_PGCB_Reverse)
summary(Rarefaction_BT_PGCB_Reverse)
attach(Rarefaction_BT_PGCB_Reverse)
names(Rarefaction_BT_PGCB_Reverse)

#ci=2 means that the CI is 95%
sp_BT <- specaccum(Rarefaction_BT_PGCB_Reverse, object="prey taxa", permutations = 100, ci=2)
#View(sp_BT)

plot(sp_BT, ylab="Number of Prey Taxa", xlab="Number of samples",col="black",ci.type="poly", ci.col="grey", ci.lty = 0,
     xlim=c(1, 32), ylim=c(1, 20)) 
```
```{r}
#Plot all samples together
#Need to highlight both lines and run them together with command return
#ci=2 means that the CI is 95%
sp_SH <- specaccum(Rarefaction_SH_Reverse, object="prey taxa", permutations = 100, ci=2)
sp_BT <- specaccum(Rarefaction_BT_PGCB_Reverse, object="prey taxa", permutations = 100, ci=2)


#Need to highlight both lines and run them together with command return
plot(sp_SH, ylab="Number of Prey Taxa", xlab="Number of samples",col="blue",ci.type="poly", ci.col="lightblue", ci.lty = 0,
     xlim=c(1, 32), ylim=c(1, 20))  
lines(sp_BT,col="darkgreen", ci.type="poly", ci.col=(rgb(0,1,0, alpha=0.3)), ci.lty = 0)
```